animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation

### 2025-07-11 animbook-journal.Rmd
fixed render arguments in yaml

---
title: "animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation"
abstract: >
  Performance is often reported in quantiles or ranks and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article "Extensive data show punishing reach of racism for black boys," we provide new visualization functions to generate animations that display this type of data in an R package called animbook. Functions to pre-process the data and prepare it for the animation are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
author:  
  - name: Krisanat Anukarnsakulchularp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    - Melbourne, Australia
    orcid: 0009-0008-5638-7124
    email:  kanu0003@student.monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: 
      - Department of Econometrics and Business Statistics
      - Melbourne, Australia
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
  rjtools::rjournal_pdf_article:
    toc: no
bibliography: animbook-journal.bib
---

@@

Two real-world examples of usage are provided. The Osiris [@bvd] example studies company movement between quartiles of sales performance over a 12-year period. The voting example studies party preference changes between the 2016 and 2019 elections.

>>>>

---
title: "animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation"
abstract: >
  Performance is often reported in quantiles or ranks and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article "Extensive data show punishing reach of racism for black boys," we provide new visualization functions to generate animations that display this type of data in an R package called animbook. Functions to pre-process the data and prepare it for the animation are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
author:  
  - name: Krisanat Anukarnsakulchularp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    - Melbourne, Australia
    orcid: 0009-0008-5638-7124
    email:  kanu0003@student.monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: 
      - Department of Econometrics and Business Statistics
      - Melbourne, Australia
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
type: package
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: animbook-journal.bib
---

@@

Two real-world examples of usage are provided. The Osiris data [@bvd] example studies company movement between quartiles of sales performance over a 12-year period. The voting example studies party preference changes between the 2016 and 2019 elections.

### 2025-07-11 animbook-journal.tex
fixed render arguments in yaml

\caption{Screenshot of the New York Times animation, which is the motivation for this visualization package. The animation illustrates the disparity between white men and black men. It reveals that despite black boys growing up in a rich household, they still have a higher likelihood of falling into lower income groups compared to white boys.}\label{fig:nyt}
\end{figure}

@@

In the data structure, there are requirements that must be followed for reproducing the animation. First, the dataset needs to be in a \texttt{tidy\ data} format (\citet{tidy-data}). The data then must have at least ID and time variables, in addition to the measured variables, which would usually be numeric but can also be categorical as well. The ID variable indicates the individual, which is followed over time, such as the company name. There may also be a grouping variable, such as the country where the company is registered (United States and Japan).

@@

\citet{mayer_2005} states that designing multimedia requires the designer to understand how people learn. One of the principles in \citet{mayer_2005}, Redundancy, suggests that a piece of excess information could overload the learners. By this principle, the animation must be carefully designed to avoid this pitfall.

@@

The \texttt{wallaby\_data()} function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function is also responsible for interpolating the non-linear path. It performs this task by mapping the \texttt{sine()} function provided in this package to each observation using the \texttt{map()} function from the \CRANpkg{purrr} package (\citet{purrr}). The frame is recalculated using the same method mentioned in the data preprocessing section. The \texttt{proportional\_shade()} function is called to generate the proportional shaded data.

The \texttt{proportional\_shade()} function's algorithm begins by calculating the corner points for each shaded area. Subsequently, the algorithm utilizes the \texttt{sine()} function to interpolate points between the left and right boundaries. Once this interpolation is completed, all the generated points are interconnected to form the shaded regions.

@@

Now that the data are in the right format for the wallaby's plot, the \texttt{geom\_point()} function is used for plotting the observations, \texttt{geom\_polygon()} is used for creating the proportional shaded areas, and \texttt{geom\_text()} is used for creating labels. These three functions are from the \CRANpkg{ggplot2} package. During this process, the aesthetics mapping will be different depending on the rendering tool to be used, which can be either \texttt{gganimate} or \texttt{plotly}. The difference between the two rendering tools is that for \texttt{plotly}, the \texttt{ids} and \texttt{frame} arguments need to be specified during the creation of the \texttt{ggplot} object.

@@

Two real-world examples of usage are provided. The osiris example studies company movement between quartiles of sales performance over a 12-year period. The voting example studies party preference changes between the 2016 and 2019 elections.

@@

The Osiris data presented in this section was acquired from Bureau van Dijk (\citet{bvd}), compiling information on a total of 122,318 companies. It was subsequently subsetted for the 2021 report, consisting of 1000 companies with 94 variables. The dataset included in the package focuses on 790 companies, spanning from 2006 to 2018. The variables in this dataset include year, ID, country, sales, and japan.

>>>>

\caption{Screenshot of the New York Times animation, which is the motivation for this visualization package. The animation illustrates the disparity between white men (yellow) and black men (blue). It reveals that despite black boys growing up in a rich household, they still have a higher likelihood of falling into lower income groups compared to white boys.This is observed by more yellow points apparently falling from the top bar to lower bars in the animated plot.}\label{fig:nyt}
\end{figure}

@@

In the data structure, there are requirements that must be followed for reproducing the animation. First, the dataset needs to be in a \texttt{tidy\ data} format (\citet{tidy-data}). The data then must have at least ID and time variables, in addition to the measured variables, which would usually be numeric but can also be categorical as well. The ID variable indicates the individual, which is followed over time, such as the company name. There may also be a grouping variable, such as the country where the company is registered (United States and Japan), which would be used to color points for group comparison.

@@

\citet{mayer_2005} states that designing multimedia requires the designer to understand how people learn. One of the principles in \citet{mayer_2005}, redundancy, suggests that a piece of excess information could overload the learners. By this principle, the animation must be carefully designed to avoid this pitfall.

@@

The \texttt{wallaby\_data()} function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function is also responsible for interpolating the non-linear path. It performs this task by mapping the \texttt{sine()} function provided in this package to each observation using the \texttt{map()} function from the \CRANpkg{purrr} package (\citet{purrr}). The frame is recalculated using the same method mentioned in the data preprocessing section.

The \texttt{proportional\_shade()} function is called to generate the proportional shaded data. The \texttt{proportional\_shade()} function's algorithm begins by calculating the corner points for each shaded area. Subsequently, the algorithm utilizes the \texttt{sine()} function to interpolate points between the left and right boundaries. Once this interpolation is completed, all the generated points are interconnected to form the shaded regions.

@@

Now that the data are in the right format for the wallaby plot, the \texttt{geom\_point()} function is used for plotting the observations, \texttt{geom\_polygon()} is used for creating the proportional shaded areas, and \texttt{geom\_text()} is used for creating labels. These three functions are from the \CRANpkg{ggplot2} package. During this process, the aesthetics mapping will be different depending on the rendering tool to be used, which can be either \texttt{gganimate} or \texttt{plotly}. The difference between the two rendering tools is that for \texttt{plotly}, the \texttt{ids} and \texttt{frame} arguments need to be specified during the creation of the \texttt{ggplot} object.

@@

Two real-world examples of usage are provided. The Osiris data \citep{bvd} example studies company movement between quartiles of sales performance over a 12-year period. The voting example studies party preference changes between the 2016 and 2019 elections.

@@

The full Osiris provides information on a total of 122,318 companies. The dataset included in the package focuses on 790 companies, spanning from 2006 to 2018. The variables in this dataset include year, ID, country, sales, and the binary variable called japan.
